Tabular Constraint Learning

### 2016-04-10 Paper/ECAI/tabular.tex
Corrections and case study evaluation

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}

@@

@@

\begin{table*}
  \centering
  \begin{tabular}{ccc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
    ALLDIFFERENT(X) & $\textual(G_X) \vee \integer(G_X) $ & $X \in G_X{:}~\forall i,j \in \nat{:}~i \neq j \rightarrow X[i] \neq X[j]$ \\
    X = Y & & \\
    FK -> PK & & \\
    R = PRODUCT(FV, FK=PK | PV) & & \\
    FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
    FV = LOOKUP(FK, PK, PV) & & \\
    PERMUTATION(X) & & \\
    R = O1 * O2 & & \\
    R = PROJECT(P) & & \\
     Y = RANK(X)     & $\numeric(G_X) \wedge \integer(G_Y) \wedge \length(G_X) = \length(G_Y)$ & $\begin{array}{r} X \in G_X, Y \in G_Y{:}~\forall i,j \in \nat{:}~(Y_i \leq Y_j \rightarrow X_{Y_i} \leq X_{Y_j}) \\ \wedge~ 1 \leq Y_i \leq \length(G_Y) \end{array}$ \\
    A = PREV(A) + P - N & & \\
    SERIES(X) & & \\
    Y = SUM(X, col) & & \\
    Y = SUM(X, row) & & \\
    R = SUMIF(FK=PK, V) & & \\
    R = SUMPRODUCT(O1, O2) & & \\

@@

\section{Case Study aka Experiments}

@@

>>>>

\documentclass{ecai}
\usepackage{times}
\usepackage{graphicx}
\usepackage{latexsym}
\usepackage{xspace}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage[numbers]{natbib}
\usepackage{notoccite}
\usepackage{framed}
\usepackage{amsmath}

@@

\newcommand{\range}[3]{\ensuremath{#1[#2,#3]}}
\newcommand{\rangeto}[2]{#1{:}#2}
\newcommand{\rangeall}{:}

\newcommand{\ecrank}[2]{\ensuremath{#1 = \mathit{RANK}(#2)}}

@@

\begin{table*}
  \centering
  \begin{tabular}{lcc}
    \textbf{\CName} & \textbf{\CSignature} & \textbf{\CFunction}\\
    ALLDIFFERENT(X) & $\textual(G_X) \vee \integer(G_X) $ & $X \in G_X{:}~\forall i,j \in \nat{:}~i \neq j \rightarrow X[i] \neq X[j]$ \\
    X = Y & & \\
    FK -> PK & & \\
    R = PRODUCT(FV, FK=PK | PV) & & \\
    FV = FUZZY-LOOKUP(FK, PK, PV) & & \\
    FV = LOOKUP(FK, PK, PV) & & \\
    PERMUTATION(X) & & \\
    R = O1 * O2 & & \\
    R = PROJECT(P) & & \\
     Y = RANK(X)     & $\numeric(G_X) \wedge \integer(G_Y) \wedge \length(G_X) = \length(G_Y)$ & $\begin{array}{r} X \in G_X, Y \in G_Y{:}~\forall i,j \in \nat{:}~(Y_i \leq Y_j \rightarrow X_{Y_i} \leq X_{Y_j}) \\ \wedge~ 1 \leq Y_i \leq \length(G_Y) \end{array}$ \\
    A = PREV(A) + P - N & & \\
    SERIES(X) & & \\
    Y = SUM(X, col) & & \\
    Y = SUM(X, row) & & \\
    R = SUMIF(FK=PK, V) & & \\
    R = SUMPRODUCT(O1, O2) & & \\

@@

@@

\section{Case Study aka Experiments}
In this section we experimentally validate our approach.
The implementation is first illustrated using a case study on a specific Excel sheet.
In order to quantify the results and generalize our findings we created a benchmark of 30 (\samuel{check number}) Excel sheets from various sources.

\subsection{Benchmark}
There are three main categories of spreadsheets in the benchmark: spreadsheets from an exercise session teaching Excel at an affiliated University, spreadsheets from tutorials online, publicly available spreadsheets such as crime statistics or financial reports\footnote{\samuel{link to links.txt}}.

All spreadsheets have been converted to CSV with no manual intervention unless noted.
Definitions of tables where added manually, providing a way to compare results and overcoming shortcomings in the table detection algorithm (in some harder cases groups where also provided manually).

For every spreadsheet a ground-truth has been provided manually, specifying the essential (or original) constraints that are expressible in Excel and expected to be discovered.
\footnote{Constraints currently outside of the scope of this system (e.g. generic nested mathematical formulas) are in a separate category (non-trivial).}

\subsection{Case study}
Let us illustrate our implementation using the example presented in the introduction.
We obtained this example by combining several Excel sheets from exercise sessions into one.
Figure~\ref{fig:sol_example} shows the constraints that we expect to find.

\paragraph{Accuracy}
Our current implementation finds 18 constraints, containing all of the 12 solutions described in figure~\ref{fig:sol_example}.
Of the 6 remaining (redundant) constraints, 5 are $\mathit{RANK}$ constraints that are true by accident, such as: \begin{align*}
  & \ecrank{\range{T_1}{\rangeall}{1}}{\range{T_1}{\rangeall}{5})} \\
  & \ecrank{\range{T_1}{\rangeall}{8}}{\range{T_1}{\rangeall}{4}}
\end{align*}
The other redundant constraint is a additional $\mathit{LOOKUP}$ that holds because the vectors in \range{T_2}{\rangeall}{\rangeto{2}{3}} can both be used to look each other up in \range{T_1}{\rangeall}{\rangeto{1}{2}} and we considered only one of them to be essential (lookup using the ID).

We conclude that for this example our primary goal of finding all constraints is achieved.
A number of redundant constraints ($33.33\%$ of the total) are also found, which, as we will show, is rather high compared to other spreadsheets.
However, this example also contains many short vectors which increases the chance for constraints to be true by accident.

\begin{figure}
  \label{fig:sol_example}
  {\small
    \begin{align*}
      & SERIES(\range{T_{1}}{\rangeall}{1}) \\
      & \range{T_{1}}{\rangeall}{8} = RANK(\range{T_{1}}{\rangeall}{7}) \\
      & \range{T_{2}}{1}{\rangeall} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
      & \range{T_{6}}{\rangeall}{2} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{6}}, col) \\
      & \range{T_{1}}{\rangeall}{7} = SUM(\range{T_{1}}{\rangeall}{\rangeto{3}{6}}, row) \\
      & \range{T_{2}}{2}{\rangeall} = AVERAGE(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
      & \range{T_{2}}{3}{\rangeall} = MAX(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
      & \range{T_{2}}{4}{\rangeall} = MIN(\range{T_{1}}{\rangeall}{\rangeto{3}{7}}, col) \\
      & \range{T_{1}}{\rangeall}{10} = SUMIF(\range{T_{5}}{\rangeall}{1}=\range{T_{1}}{\rangeall}{2}, \range{T_{5}}{\rangeall}{2}) \\
      & \range{T_{1}}{\rangeall}{11} = MAXIF(\range{T_{5}}{\rangeall}{1}=\range{T_{1}}{\rangeall}{2}, \range{T_{5}}{\rangeall}{2}) \\
      & \range{T_{4}}{\rangeall}{3} = LOOKUP(\range{T_{4}}{\rangeall}{2}, \range{T_{1}}{\rangeall}{1}, \range{T_{1}}{\rangeall}{2}) \\
      & \range{T_{6}}{\rangeall}{4} = PREV(\range{T_{6}}{\rangeall}{4}) + \range{T_{6}}{\rangeall}{2} - \range{T_{6}}{\rangeall}{3}
    \end{align*}
  }%
  \caption{Expected constraints case study}
\end{figure}
